---
format:
  html:
    css: styles.css
execute:
  echo: false
---

<style>
body {
  background: linear-gradient(135deg, #0b0c1a, #1a0533, #0d0221, #0b1a2e);
  background-size: 400% 400%;
  animation: gradientShift 12s ease infinite;
  min-height: 100vh;
}

@keyframes gradientShift {
  0%   { background-position: 0% 50%; }
  50%  { background-position: 100% 50%; }
  100% { background-position: 0% 50%; }
}

.reveal {
  opacity: 0;
  transform: translateY(36px);
  transition: opacity 0.75s cubic-bezier(.22,.68,0,1.2), transform 0.75s cubic-bezier(.22,.68,0,1.2);
  will-change: opacity, transform;
}
.reveal.is-visible {
  opacity: 1;
  transform: translateY(0);
}

.chart-fade {
  opacity: 0;
  transform: scale(0.97);
  transition: opacity 0.9s ease 0.3s, transform 0.9s ease 0.3s;
}
.chart-fade.is-visible {
  opacity: 1;
  transform: scale(1);
}

.hero-panel {
  display: flex;
  gap: 2rem;
  align-items: center;
  padding: 3rem 2rem;
  border-radius: 16px;
  margin-bottom: 2rem;
  background: rgba(255,255,255,0.04);
  border: 1px solid rgba(255,255,255,0.08);
  backdrop-filter: blur(6px);
}

.hero-copy { flex: 1; }

.eyebrow {
  font-family: monospace;
  font-size: 11px;
  letter-spacing: 0.15em;
  text-transform: uppercase;
  color: #9b6dff;
  margin-bottom: 0.6rem;
}

.hero-copy h1 {
  font-size: clamp(1.6rem, 3vw, 2.4rem);
  font-weight: 800;
  color: #f0eeff;
  line-height: 1.2;
  margin: 0 0 0.75rem 0;
}

.hero-copy p {
  color: #a0a0c8;
  font-family: monospace;
  font-size: 14px;
  line-height: 1.65;
  max-width: 480px;
}

.hero-pills {
  display: flex;
  gap: 0.5rem;
  flex-wrap: wrap;
  margin-top: 1.25rem;
}

.hero-pill {
  font-family: monospace;
  font-size: 12px;
  padding: 4px 12px;
  border-radius: 999px;
  background: rgba(155, 109, 255, 0.15);
  border: 1px solid rgba(155, 109, 255, 0.35);
  color: #c9aaff;
  transition: background 0.2s, transform 0.2s;
}

.hero-pill:hover {
  background: rgba(155, 109, 255, 0.3);
  transform: translateY(-2px);
}

.hero-card {
  display: flex;
  flex-direction: column;
  gap: 1rem;
  min-width: 200px;
}

.hero-metric {
  background: rgba(255,255,255,0.05);
  border: 1px solid rgba(255,255,255,0.1);
  border-radius: 12px;
  padding: 1rem 1.25rem;
  transition: transform 0.25s ease, box-shadow 0.25s ease, background 0.25s ease;
  cursor: default;
}

.hero-metric:hover {
  transform: translateY(-4px) scale(1.03);
  box-shadow: 0 8px 30px rgba(155, 109, 255, 0.25);
  background: rgba(155, 109, 255, 0.12);
}

.hero-metric strong {
  display: block;
  font-size: 1.8rem;
  font-weight: 800;
  color: #fffb96;
  font-family: monospace;
  line-height: 1;
}

.hero-metric span {
  font-family: monospace;
  font-size: 11px;
  color: #8888aa;
  text-transform: uppercase;
  letter-spacing: 0.08em;
}

.counter-val {
  display: inline-block;
  transition: opacity 0.2s;
}

.panel {
  background: rgba(255,255,255,0.04);
  border: 1px solid rgba(255,255,255,0.08);
  border-radius: 12px;
  padding: 1.5rem 2rem;
  margin: 1.5rem 0;
  transition: transform 0.25s ease, box-shadow 0.25s ease;
}

.panel:hover {
  transform: translateY(-3px);
  box-shadow: 0 6px 24px rgba(100, 80, 200, 0.2);
}

.panel h3 {
  color: #d0c8ff;
  font-family: monospace;
  margin-top: 0;
}

.panel p {
  color: #9090b8;
  font-family: monospace;
  font-size: 14px;
  line-height: 1.6;
  margin-bottom: 0;
}
</style>

```{r}
#| message: false
#| warning: false
#| include: false

library(tidyverse)
library(plotly)
library(htmltools)
library(reactable)

games <- read_csv("Games.csv")

games <- games |>
  mutate(
    gameDate = as.Date(gameDate),
    result = case_when(
      homeScore > awayScore ~ "Home win",
      homeScore < awayScore ~ "Away win",
      TRUE ~ "Tie"
    )
  )

score_counts <- games |>
  select(home = homeScore, away = awayScore,
         homeTeam = hometeamName, awayTeam = awayteamName,
         date = gameDate, result) |>
  drop_na() |>
  filter(home >= 50, away >= 50, home <= 200, away <= 200) |>
  mutate(
    date = as.Date(date),
    matchup = paste0(format(date, "%b %d, %Y"), " — ", homeTeam, " vs ", awayTeam)
  ) |>
  group_by(home, away, result) |>
  summarise(
    n_games = n(),
    matchups = {
      u <- sort(unique(matchup))
      if (length(u) > 10) paste0(paste(u[1:10], collapse = "<br>"), "<br>… and ", length(u) - 10, " more")
      else paste(u, collapse = "<br>")
    },
    .groups = "drop"
  ) |>
  mutate(
    hover = paste0(
      "<b>Home ", home, "  –  Away ", away, "</b><br>",
      "Result: ", result, "<br>",
      "Times occurred: ", n_games, "<br><br>",
      matchups
    )
  )

n_unique <- nrow(score_counts)
year_min <- format(min(games$gameDate, na.rm = TRUE), "%Y")
year_max <- format(max(games$gameDate, na.rm = TRUE), "%Y")
```

::: {.hero-panel .reveal}
::: hero-copy
::: eyebrow
Interactive Quarto dashboard
:::

# Explore NBA scorelines like a live game study. {.unnumbered}

Hover across the scatter plot, toggle result views, and browse the most common team matchups in a polished, animated experience.

::: hero-pills
[✨ Scroll reveals]{.hero-pill}
[📈 Live chart controls]{.hero-pill}
[🔎 Searchable table]{.hero-pill}
:::
:::

::: hero-card
::: {.hero-metric .js-animated}
**[100+]{.counter-val data-target="100" data-suffix="+"}**

[unique scoring patterns visible]{style="font-family:monospace;font-size:11px;color:#8888aa;text-transform:uppercase;letter-spacing:0.08em;"}
:::

::: {.hero-metric .js-animated}
**[3]{.counter-val data-target="3" data-suffix=""}**

[interactive view modes]{style="font-family:monospace;font-size:11px;color:#8888aa;text-transform:uppercase;letter-spacing:0.08em;"}
:::

::: {.hero-metric .js-animated}
**[Top 10]{.counter-val data-target="10" data-prefix="Top " data-suffix=""}**

[teams ranked by appearances]{style="font-family:monospace;font-size:11px;color:#8888aa;text-transform:uppercase;letter-spacing:0.08em;"}
:::
:::
:::

::: {.panel .reveal}
### Interactive NBA score exploration

This Quarto site turns the NBA game data into a live dashboard. Hover over the points, switch views with the buttons, and explore the most frequently appearing teams in the table below.
:::

::: {.chart-fade .reveal}
```{r}
#| message: false
#| warning: false

fig <- plot_ly() |>
  add_trace(
    data = score_counts,
    x = ~home,
    y = ~away,
    type = "scatter",
    mode = "markers",
    name = "All games",
    visible = TRUE,
    marker = list(
      color = ~log1p(n_games),
      colorscale = list(
        list(0, "#0d0221"),
        list(0.2, "#1a1a5e"),
        list(0.4, "#6a0572"),
        list(0.6, "#c0392b"),
        list(0.8, "#f39c12"),
        list(1, "#fffb96")
      ),
      cmin = 0,
      cmax = log1p(max(score_counts$n_games)),
      size = 7,
      opacity = 1,
      line = list(width = 0)
    ),
    text = ~hover,
    hoverinfo = "text"
  ) |>
  add_trace(
    data = score_counts |> filter(result == "Home win"),
    x = ~home,
    y = ~away,
    type = "scatter",
    mode = "markers",
    name = "Home wins",
    visible = FALSE,
    marker = list(
      color = ~log1p(n_games),
      colorscale = list(
        list(0, "#0d0221"),
        list(0.2, "#1a1a5e"),
        list(0.4, "#6a0572"),
        list(0.6, "#c0392b"),
        list(0.8, "#f39c12"),
        list(1, "#fffb96")
      ),
      cmin = 0,
      cmax = log1p(max(score_counts$n_games)),
      size = 7,
      opacity = 1,
      line = list(width = 0)
    ),
    text = ~hover,
    hoverinfo = "text"
  ) |>
  add_trace(
    data = score_counts |> filter(result == "Away win"),
    x = ~home,
    y = ~away,
    type = "scatter",
    mode = "markers",
    name = "Away wins",
    visible = FALSE,
    marker = list(
      color = ~log1p(n_games),
      colorscale = list(
        list(0, "#0d0221"),
        list(0.2, "#1a1a5e"),
        list(0.4, "#6a0572"),
        list(0.6, "#c0392b"),
        list(0.8, "#f39c12"),
        list(1, "#fffb96")
      ),
      cmin = 0,
      cmax = log1p(max(score_counts$n_games)),
      size = 7,
      opacity = 1,
      line = list(width = 0)
    ),
    text = ~hover,
    hoverinfo = "text"
  ) |>
  add_segments(
    x = 60, xend = 190, y = 60, yend = 190,
    line = list(color = "rgba(0,0,0,0.15)", width = 1, dash = "dash"),
    hoverinfo = "skip",
    showlegend = FALSE,
    inherit = FALSE
  ) |>
  layout(
    paper_bgcolor = "#ffffff",
    plot_bgcolor = "#f8f9fa",
    font = list(color = "#333333", family = "monospace"),
    title = list(text = "NBA Scorigami", x = 0.5),
    xaxis = list(
      title = list(text = "Home PTS", font = list(color = "#555555", size = 16)),
      tickfont = list(color = "#666666", size = 14),
      tickangle = 0,
      gridcolor = "#e0e0e0",
      zerolinecolor = "#d0d0d0",
      dtick = 10,
      range = c(58, 195),
      fixedrange = FALSE
    ),
    yaxis = list(
      title = list(text = "Visitor PTS", font = list(color = "#555555", size = 16)),
      tickfont = list(color = "#666666", size = 14),
      gridcolor = "#e0e0e0",
      zerolinecolor = "#d0d0d0",
      dtick = 10,
      range = c(58, 195),
      scaleanchor = "x",
      scaleratio = 1,
      fixedrange = FALSE
    ),
    margin = list(t = 120, r = 50, b = 100, l = 100),
    height = 700,
    hoverlabel = list(
      bgcolor = "#f0f0f5",
      bordercolor = "#999999",
      font = list(color = "#333333", size = 12, family = "monospace")
    ),
    updatemenus = list(
      list(
        type = "buttons",
        direction = "right",
        x = 0.5,
        y = 1.11,
        xanchor = "center",
        yanchor = "top",
        bgcolor = "#f8f9fa",
        bordercolor = "#d0d0d0",
        showactive = TRUE,
        buttons = list(
          list(
            label = "All games",
            method = "update",
            args = list(list(visible = c(TRUE, FALSE, FALSE)), list(title = "All games"))
          ),
          list(
            label = "Home wins",
            method = "update",
            args = list(list(visible = c(FALSE, TRUE, FALSE)), list(title = "Home wins"))
          ),
          list(
            label = "Away wins",
            method = "update",
            args = list(list(visible = c(FALSE, FALSE, TRUE)), list(title = "Away wins"))
          )
        )
      )
    )
  ) |>
  config(
    displayModeBar = TRUE,
    modeBarButtonsToRemove = c("select2d", "lasso2d", "autoScale2d"),
    displaylogo = FALSE,
    scrollZoom = TRUE
  )

fig
```
:::

::: {.panel .reveal}
### Scoreline highlights

These points reveal where final totals cluster, making it easier to see how often specific results repeat across seasons.
:::

```{r}
#| message: false
#| warning: false

htmltools::div(
  style = "margin: 30px 0 40px 0; text-align: center; color: #888888; font-family: monospace; font-size: 12px;",
  paste0("Data: NBA games from ", year_min, " to ", year_max, " | Hover over dots for details | ", format(n_unique, big.mark = ","), " unique scorelines")
)
```

```{r}
#| message: false
#| warning: false

top10 <- games |>
  pivot_longer(
    cols = c(hometeamName, awayteamName),
    names_to = "side",
    values_to = "team"
  ) |>
  count(team, name = "games_played") |>
  arrange(desc(games_played)) |>
  slice_head(n = 10) |>
  mutate(rank = row_number(), .before = team)

tbl <- reactable(
  top10,
  searchable = TRUE,
  filterable = TRUE,
  defaultPageSize = 10,
  paginationType = "simple",
  columns = list(
    rank = colDef(name = "#", maxWidth = 45,
                 style = list(color = "#6666aa", fontWeight = "bold")),
    team = colDef(name = "Team", minWidth = 160),
    games_played = colDef(name = "Games Played", minWidth = 120,
                          align = "right",
                          format = colFormat(separators = TRUE),
                          cell = function(value) {
                            max_val <- max(top10$games_played)
                            pct <- value / max_val
                            htmltools::div(
                              style = "display:flex; align-items:center; gap:8px; justify-content:flex-end;",
                              htmltools::div(
                                style = paste0(
                                  "background:#4b0082; width:", round(pct * 120), "px;",
                                  "height:8px; border-radius:4px;",
                                  "background: linear-gradient(90deg, #4b0082, #e84545);"
                                )
                              ),
                              htmltools::span(format(value, big.mark = ","), style = "color:#e0e0ff; min-width:40px;")
                            )
                          })
  ),
  theme = reactableTheme(
    backgroundColor = "#0b0c1a",
    borderColor = "#1e1f33",
    stripedColor = "#10112a",
    highlightColor = "#1a1b3a",
    color = "#d0d0f0",
    headerStyle = list(
      background = "#0d0e22",
      color = "#8888cc",
      fontFamily = "monospace",
      fontSize = "12px",
      borderColor = "#1e1f33"
    ),
    style = list(fontFamily = "monospace", fontSize = "14px")
  ),
  striped = TRUE,
  highlight = TRUE,
  borderless = FALSE
)

browsable(
  tags$div(
    style = "margin-top: 40px;",
    tbl
  )
)
```

::: {.panel .reveal}
### Browse the leaderboard

Search teams, filter the table, and discover which franchises appear most often in the dataset.
:::

<script>
  document.addEventListener('DOMContentLoaded', () => {

    const revealItems = document.querySelectorAll('.reveal');
    const observer = new IntersectionObserver((entries) => {
      entries.forEach((entry) => {
        if (entry.isIntersecting) {
          entry.target.classList.add('is-visible');
          if (entry.target.classList.contains('hero-panel')) {
            startCounters();
          }
        }
      });
    }, { threshold: 0.15 });

    revealItems.forEach((item) => observer.observe(item));

    const chartFades = document.querySelectorAll('.chart-fade');
    const chartObserver = new IntersectionObserver((entries) => {
      entries.forEach((entry) => {
        if (entry.isIntersecting) {
          entry.target.classList.add('is-visible');
        }
      });
    }, { threshold: 0.05 });

    chartFades.forEach((el) => chartObserver.observe(el));

    const cards = document.querySelectorAll('.js-animated');
    cards.forEach((card, index) => {
      card.style.transitionDelay = `${index * 90}ms`;
    });

    let countersStarted = false;

    function animateCounter(el) {
      const target = parseInt(el.dataset.target, 10);
      const prefix = el.dataset.prefix || '';
      const suffix = el.dataset.suffix || '';
      const duration = 1200;
      const startTime = performance.now();

      function tick(now) {
        const elapsed = now - startTime;
        const progress = Math.min(elapsed / duration, 1);
        const eased = 1 - Math.pow(1 - progress, 3);
        const current = Math.floor(eased * target);
        el.textContent = prefix + current + suffix;
        if (progress < 1) requestAnimationFrame(tick);
      }

      requestAnimationFrame(tick);
    }

    function startCounters() {
      if (countersStarted) return;
      countersStarted = true;
      document.querySelectorAll('.counter-val').forEach((el) => {
        setTimeout(() => animateCounter(el), 300);
      });
    }

    setTimeout(() => {
      const hero = document.querySelector('.hero-panel');
      if (hero && hero.classList.contains('is-visible')) startCounters();
    }, 500);

  });
</script>
